Edge sign prediction in general graphs and Correlation Clustering

### 2017-11-07 edge/related.tex
remove draft text on edge related

\Textcite{Taskar04} represent a node-attributed graphs with partially labelled edges as a
probabilistic graphical model (called Relational Markov Model) and learns its parameters from the
data using gradient descent. It is a flexible modeling approach that let the designer express
complex graph patterns. Using a slightly less principled approach, \textcite{conceptualLinks12}
first cluster nodes based solely on their attributes and then count the links between such groups to
keep only the maximally frequent ones, arguing this reveals the set of attributes that support the
connection between two nodes. In the same data mining vein, and related to our departure from global
homophily, \textcite{beyondHome16} look for link patterns whose support and confidence is not only
above some given thresholds, but that also diverge from homophily. In the author explicitly exclude
attributes taking the same value within such link patterns.

Groups of nodes are first created when they have common attributes. The set of links between two
groups is called a conceptual link in the sense that it corresponds to a relationship between two
sets of attributes, namely two concepts in the field of formal concept analysis [7]. The number of
links between two groups is then evaluated and when the frequency is greater than a given support
threshold $\beta$, we talk about frequent conceptual link. Note that we only keep maximal frequent
conceptual links (mfcl), i.e. those who are not included in others.  Finally, the set of maximal
frequent conceptual links is used to create a new network structure called a conceptual view which
summarizes all the knowledge extracted from the initial network. In a conceptual view, each node
corresponds to a set of attributes (in this context we call them Meta-nodes) and a link corresponds
to maximal frequent conceptual links~\autocite{conceptualLinks12}.
The objective of \textcite{ZhangModelFree16} is eventually to hard cluster the nodes into
communities but their work similarities with ours. Namely, they tweak the traditional modularity
objective (maximizing the density of intra community edges) by weighting the edges with the
similarity of their endpoints, and simultaneously learning a weight vector $w_\ell$ for a community
$C_\ell$. Formally they defined the $R$ criterion as :
\begin{equation*}
  R(\mathcal{C}, \beta;\alpha, \mu) = \sum_{\ell=1}^k \frac{1}{|\mathcal{C}_\ell|^\alpha}
  \sum_{u,v \in \mathcal{C}_\ell} A_{uv} \mu -e^{-g(s_{uv}, w_\ell)}
\end{equation*}
where $A$ is the adjacency matrix, $\alpha$ penalizes unbalanced communities, $\mu$ is a trade off
parameter between information from edges and node features, and $g$ is exactly our goodness
function.  They optimize it by alternately optimizing over the labels with fixed parameters and over
the parameters with fixed labels, using block coordinate descent.
Finally we mention a more quantitative work. \Textcite{Abraham2012a} assume that there are $K$
social categories modelled by $K$ Euclidean spaces $\mathcal{D}_i$. Nodes of the graph have an
associated point in each of these spaces. The key assumption is that space have small local
correlation: informally, the intersection of any two small balls from two distinct spaces is small.
These $K$ spaces give rise to small world networks $\mathcal{G}_i$, where the edge probability
$\propto \mathcal{D}_i(u, v)^{-d}$ and we observe the real network $\mathcal{G} =\bigcup_i
\mathcal{G}_i$. From $\mathcal{G}$, the proposed algorithm recovers in $n \mathrm{polylog} n$ time a
bounded approximation $\mathcal{D}'_i$ of all $\mathcal{D}_i$, that is
\begin{equation*}
  \sigma \mathcal{D}_i(u, v) \leq \mathcal{D}'_i(u, v) \leq \delta \mathcal{D}_i(u, v) + \Delta
\end{equation*}

\iffalse
adhoc~\autocites{conceptualLinks12}{Abraham2012a}{Taskar04}.
A supervised probabilistic model
find a partition maximizing the density of intra community edges ---weighted by the endpoints
profile similarities and a learned community weight vector~\autocite{ZhangModelFree16}

Let $A$ be the unweighted adjacency matrix and  $\mathcal{E} = \bigcup_{i=1}^K
\mathcal{E}_k$ be a partition of the node, which alternatively is represented by
the community memberships vector $e$.  They start by presenting a simple
community goodness criterion, which like modularity basically score a partition
based on the number (or density) of edges within communities
$$ R(e;\alpha) = \sum_{k=1}^K \frac{1}{|\mathcal{E}_k|^\alpha} \sum_{i,j\in \mathcal{E}_k}A_{ij} $$
where $\alpha$ is a tuning parameters which penalizes unbalanced communities as
it gets larger.
The idea would be to add weight on the edges based on the node features
$f_i\in \mathbb{R}^p$ so as to increase that score. Because features and
communities don't correlate perfectly, they introduce a more flexible criterion
$$ R(e,\beta;\alpha,w_n) = \sum_{k=1}^K \frac{1}{|\mathcal{E}_k|^\alpha}
\sum_{i,j\in \mathcal{E}_k}A_{ij}W(f_i, f_j,\beta_k;w_n) $$
where
$\beta_k\in \mathbb{R}^p$ is a weight vector for the $k$th community and $w_n$
is trade off parameters between information from edges and node features.
Basically, $W(f_i, f_j,\beta_k;w_n)$ is a function of the similarity
$\phi_{ij}=\phi(f_i, f_j)$ between two nodes. This reminds of our approach, by
assigning one vector $w$ to each community. As a practical example, they use
$$
w_{ijk} = W(f_i, f_j,\beta_k;w_n) = w_n -e^{-\langle \phi_{ij},\beta_k\rangle}
$$

This \enquote{criterion needs to be optimized over both the community
assignments $e$ and the feature parameters $\beta$. Using block coordinate
descent, we optimize it by alternately optimizing over the labels with fixed
parameters and over the parameters with fixed labels, and iterating until
convergence}. Again, this is similar to what we tried at first.

\fi

@@

\iffalse
\begin{itemize}[leftmargin=*]
\item \href{http://dl.acm.org/citation.cfm?id=2629616}{A General Bayesian
  Framework for Attributed Graph Clustering}~\autocite{Xu2014}

They present a generative model of clustered attributed graph. That is, given a
vertex set $V$, \enquote{a joint probability distribution over the space of all
possible attributed graphs and all possible partitions over $V$. The
clustering with the highest probability is this clustering best explains the
attribute values and edge patterns of the given graph.}

Let $X$ be the $n\times n$ adjacency matrix, $Y$ be the $n\times T$ attribute
matrix and $Z$ be the $n$-long cluster indicator vector (taking values in
${1,\ldots,k}$). The generative process is as follow, for each node $v_i\in V$:

\begin{itemize}
\item
  Choose the cluster according to a multinomial distribution:
  $Z_i \sim \mathtt{Multinomial}(\alpha)$. That is
  $P(Z_i = k|\alpha) = \alpha_k$ with $\alpha_k \in [0,1]$ and
  $\sum_k \alpha_k = 1$.
\item
  Then sample the attribute of the node, from a distribution
  parametrized according to the chosen cluster
  $Y_i \sim P(Y_i |\theta_{Z_i}) = \prod_{t=1}^{T} P(Y_{it}|\theta_{Z_{i},t})$.
  The exact form of $P$ depends of the kind of attribute being
  modelled (binary, categorical, continuous and so on).
\item
  Finally, we potentially connect the vertex $v_i$ to the previous
  ones ${v_j, j<i}$ according to
  $X_{ij} \sim P(X_{ij} |\Phi_{Z_iZ_j})$. The intuition is that two
  vertices from the same cluster will have similar connectivity pattern
  with the rest of the graph.
\end{itemize}

In Section 3.2.2, the authors also described the prior placed on $\alpha$,
$\theta$ and $\Phi$. Note that contrary to the previous model (and because the
focus is on clustering) the edge probabilities only depend indirectly of the
attributes (through the $Z$ variable).

They use a variational approach to perform the parameters inference.

TODO: add this one \fullcite{Zhao2017}

\item \href{https://cs.stanford.edu/people/jure/pubs/cesna-icdm13.pdf}{Community
  Detection in Networks with Node Attributes}~\autocite{Yang2013}

CESNA is based on the following properties:

\begin{quote}
\begin{itemize}[leftmargin=*,nosep]
\item
  ``Nodes that belong to the same communities are likely to be connected
  to each other.
\item
  Communities can overlap, as individual nodes may belong to multiple
  communities.
\item
  If two nodes belong to multiple common communities, they are more
  likely to be connected than if they share only a single common
  community (i.e., overlapping communities are denser).
\item
  Nodes in the same community are likely to share common attributes ---
  for example, a community might consist of friends attending a same
  school.''
\end{itemize}
\end{quote}

The graph $G$ has $C$ communities and $n$ nodes, each $u$ of
them having $K$ binary attributes in its profile vector $X_u$ and
belonging to a community $c$ with intensity
$F_{uc} \in [0, \infty)$.

Link generation is done by considering that the probability of linking
two nodes $u$ and $v$ belonging to the same community $c$ is
$P_{uv}(c) = 1-\exp(F_{uc}F_{vc})$ (which is $0$ is one node does
not belong to $c$). Then, assuming that each community connects $u$
and $v$ independently, we have that
$1-P_{uv} = \prod_c (1-P_{uv}(c)) = \exp(-\sum_c F_{uc}F_{vc})$ from
which $P_{uv} = 1-\exp(-\sum_c F_{uc}F_{vc})$.

\begin{minipage}[t]{0.6\linewidth}
Likewise community memberships generate attributes. The intuition is
that given its community memberships, it should be possible to predict a
node's attributes. That is, considering $F_{u1},\ldots,F_{uC}$ as
input feature of a logistic model, and given the logistic weight
$W_{kc}$ (for attribute $k$ and community $c$), we have
\[Q_{uk} = \frac{1}{1+\exp(-\sum_c W_{kc}F_{uc})}\] and
$X_{uk} \sim \mathrm{Bernoulli}(Q_{uk})$.
\end{minipage}
\begin{minipage}[t]{0.37\linewidth}
  \centering
  \vspace{-\ht\strutbox}
  \includegraphics[width=1\linewidth]{raw/cesna.png}
\end{minipage}

\item add this one (not a generative model) \fullcite{Chesnokov2017}

\item \href{https://arxiv.org/abs/1608.00920}{Community Detection Algorithm
  Combining Stochastic Block Model and Attribute Data Clustering}~\autocite{Kataoka2016}

That one is familiar. First they also made the assumption that the adjacency
matrix is conditionally independent of attribute data given the community
assignment. Then for a given node $i$, first they pick its community $x_i$ from
a multinomial parametrized by $\gamma$.  Then its (one dimensional continuous)
attribute is drawn from the normal distribution of the community $\ell$:
$\mathcal{N}(\mu_\ell, \sigma_\ell)$. Finally the edges are created by a
Stochastic Block Model.

\item \href{https://arxiv.org/abs/1610.09735}{Community Detection with Nodal
  Information}~\autocite{Weng2016}

Here $A$ is the adjacency matrix, $X$ the attribute matrix and $c$ the community
assignment. Their first assumption is that $A \bot X | c$, \enquote{a strong
constraint stating that given the community membership, what nodes are like
(described by covariates $X$) does not affect how they are connected (encoded
in $A$)}.

This motivates them to consider the following model \emph{Node-coupled
Stochastic Block Model (NSBM)}:
\begin{align*}
  P(c|X) &= \prod_i \frac{\exp( \beta^T_{c_i} x_i)}{\sum^K_{k =1}\exp( \beta^T_{c_i} x_i)} \\
  P(A|c) &= \prod_{ i<j} B^{A_{ij}}_{c_i c_j}(1 - B_{c_i c_j})^{(1-A_{ij})}
\end{align*}
where $B = ( B_{ab}) \in [0, 1]^{K\times K}$ is symmetric and
$\beta = ( \beta_1 , \ldots, \beta_K) \in \mathbb{R}^{Kp}$.

\begin{quote}
\enquote{
The distribution $P(A|c)$ follows the stochastic block model (SBM), which, as
the fundamental model, has been extensively studied in the literature. The SBM
implies that the distribution of an edge between node $i$ and $j$ only depends
on their community membership $c_i$ and $c_j$.

The nodes from the same community are stochastically equivalent. The element
$B_{ab}$ in the matrix $B$ represents the probability of edge connection between
a node in community $a$ and a node in community $b$.

The $P ( c | X)$ simply takes a multi-logistic regression form, where we will
assume $\beta_K = 0$ for identifiability.}
\end{quote}

\item\href{https://arxiv.org/abs/1507.04001}{Structure and Inference in
  Annotated Networks}~\autocite{Newman2016}

Consider $n$ nodes labeled $u=1,\ldots,n$, divided into $k$ communities ($u$
belongs to community $s_u$), and associated with one metadata $x_u \in [1, K]$
(that is a simplified version for introductory purpose, later metadata can also
be continuous). Given $\{x_u\}_{u=1}^n$ and the degree sequence
$\{d_u\}_{u=1}^n$, the network is generated as follow. First each node $u$ is
assigned to community $s$ with a probability $\gamma_{s,x_u}$ (the $k\times K$
matrix $\Gamma$ holds these parameters). One each node is assigned a community,
we use a degree corrected block model to create the network.  Edge between $u$
and $v$ is created with probability $p_{uv} = d_ud_v\theta_{s_u,s_v}$. This
time, $\theta$ parameters are stored in a $k\times k$ symmetric matrix $\Theta$.

\item \href{http://doi.org/10.1103/PhysRevE.95.042317}{\citefield{DeBacco2017}{title}}
  \autocite{DeBacco2017}

\end{itemize}

It has been implemented in many ways (make a table?)
Replace distance measure by a Bayesian model~\autocites{Xu2012}{Xu2014}{Zhao2017}.
CESNA (see maingraphSC)~\autocite{Yang2013}, overlapping, compare with CESNA so put them together
\autocite{Chesnokov2017}
generative models and then node membership inference~\autocites{Kataoka2016}{Weng2016}{Newman2016}
\fi

% keep checking in mendeley and community from maingraphSC
%
% \subsection{Mendeley}

% \begin{description}
% \item[\citefield{Zhou2009}{title}] -- \fullcite{Zhou2009}\\ \citefield{Zhou2009}{abstract}
% \item[\citefield{Zhou2010}{title}] -- \fullcite{Zhou2010}\\ \citefield{Zhou2010}{abstract}

% \item[\citefield{Xu2012}{title}] -- \fullcite{Xu2012}\\ \citefield{Xu2012}{abstract}
% \item[\citefield{Xu2014}{title}] -- \fullcite{Xu2014}\\ \citefield{Xu2014}{abstract}

% \item[\citefield{Silva2012}{title}] -- \fullcite{Silva2012}\\ \citefield{Silva2012}{abstract}

  % too far
% \item[\citefield{Ruan2013}{title}] -- \fullcite{Ruan2013}\\ \citefield{Ruan2013}{abstract}

% CESNA
% \item[\citefield{Yang2013}{title}] -- \fullcite{Yang2013}\\ \citefield{Yang2013}{abstract}

% subspace
% \item[\citefield{Huang2015}{title}] -- \fullcite{Huang2015}\\ \citefield{Huang2015}{abstract}

% Too far an not super clear, I think they do node classification
% \item[\citefield{Bayer2015}{title}] -- \fullcite{Bayer2015}\\ \citefield{Bayer2015}{abstract}

% \item[\citefield{Kataoka2016}{title}] -- \fullcite{Kataoka2016}\\ \citefield{Kataoka2016}{abstract}

% \item[\citefield{Weng2016}{title}] -- \fullcite{Weng2016}\\ \citefield{Weng2016}{abstract}

% link with k-CC
% \item[\citefield{Barnes2016a}{title}] -- \fullcite{Barnes2016a}\\ \citefield{Barnes2016a}{abstract}

% \item[\citefield{Newman2016}{title}] -- \fullcite{Newman2016}\\ \citefield{Newman2016}{abstract}

  % Unpublished, benchmarks? (Mason Porter though)
% \item[\citefield{Bazzi2016}{title}] -- \fullcite{Bazzi2016}\\ \citefield{Bazzi2016}{abstract}

  % KG
% \item[\citefield{Ren2017}{title}] -- \fullcite{Ren2017}\\ \citefield{Ren2017}{abstract}

  % fun extension, each nodes is itself a single-attribute graph…
% \item[\citefield{Guo2017}{title}] -- \fullcite{Guo2017}\\ \citefield{Guo2017}{abstract}

% MiMAG mining dense subgraphs featuring a single edge type, so it's more of downstream application
  % “we aim at finding clusters of vertices that are densely connected by edges with similar labels
  % in a subset of the graph layers” paper version of
% \item[\citefield{Boden2017}{title}] -- \fullcite{Boden2017}\\ \citefield{Boden2017}{abstract}

  % Unpublished, quite far anyway
% \item[\citefield{Wu2017}{title}] -- \fullcite{Wu2017}\\ \citefield{Wu2017}{abstract}

  % share the idea of explaining edges: “Given a collection of attributed subgraphs from different
  % classes, how can we discover the attributes that characterize their differences?” thing is, it's
  % about differences so… and there Focus attribute KDD'14 paper is only about node
% \item[\citefield{Rezaei2017}{title}] -- \fullcite{Rezaei2017}\\ \citefield{Rezaei2017}{abstract}

  % Anomaly detection is a nice application in edge-labeled graphs: Novel distributed algorithms
  % that identify problematic nodes whose influence can only be detected on their neighbors,
  % validated through the analysis of data breaches in bank transactions. (except there's no label
  % here) cite it rather for the tensor stuff
% \item[\citefield{Araujo2017}{title}] -- \fullcite{Araujo2017}\\ \citefield{Araujo2017}{abstract}

% \item[\citefield{Baroni2017}{title}] -- \fullcite{Baroni2017}\\ \citefield{Baroni2017}{abstract}

  % nice keywords in the abstract: overlapping, disassortative, layer merging for compression,
  % generative model: the model assigns probabilities to each pair of nodes that they have an
  % unobserved link of each type. However, it needs supervision
  % node have mixed membership to K groups and each layer is generated by a K×K matrix given the
  % density of edges between groups in that specific layer, taking into account the group
  % membership, like in a degree corrected SBM. Unlike us, this allows for multigraphs, but the EM
  % procedure presented to recover the model parameters require observing all the layers, that is
  % full supervision. Once its parameters are found, this model can then be used to predict the
  % existence of extra edges in each layer, which can be seen as edge type prediction. Furthermore,
  % “measuring the extent to which one layer helps predict links in another layer provide a
  % way to measure the relationships between the layers, from redundancy to independence, allowing
  % information compression.
% \item[\citefield{DeBacco2017}{title}] -- \fullcite{DeBacco2017}\\ \citefield{DeBacco2017}{abstract}

  % SNE embedding, move it to conclusion
% \item[\citefield{Liao2017}{title}] -- \fullcite{Liao2017}\\ \citefield{Liao2017}{abstract}

  % Too far, detecting localized attributes on a graph
% \item[\citefield{Chen2017}{title}] -- \fullcite{Chen2017}\\ \citefield{Chen2017}{abstract}

  %ICML, another generative model with fast, fully Bayesian inference but no edge type.
% \item[\citefield{Zhao2017}{title}] -- \fullcite{Zhao2017}\\ \citefield{Zhao2017}{abstract}

  % overlapping, compare with CESNA so put them together
% \item[\citefield{Chesnokov2017}{title}] -- \fullcite{Chesnokov2017}\\ OVERLAPPING \citefield{Chesnokov2017}{abstract}

% \end{description}

% \cite{Zhou2009,Zhou2010,Xu2012,Silva2012,Ruan2013,Yang2013,Xu2014,Huang2015,Bayer2015,BOTHOREL2015,Kataoka2016,Weng2016,Barnes2016a,Newman2016,Bazzi2016,Ren2017,Guo2017,Boden2017,Wu2017,Rezaei2017,Araujo2017,Baroni2017,DeBacco2017,Liao2017,Chen2017,Zhao2017}

\iffalse

\Textcite[section 7.2 edge labelling]{nodeClassif11} mention a cool paper from Darja Krushevskaja
but I can't find it online.

\fullcite{Zhuang2012} and
section 2.3 of Semantic Mining of Social Networks (same thing I believe.
Actually they also talk about a WSDM paper on infering ties using transfer
learning~\autocite{transferTiesPred16})

\Textcite[Section 4.2.2]{miningSocialTheoriesSurvey14} is about \emph{Social Tie
Prediction}, but they only mention \autocite{transferTiesPred16} and
\autocite{Yang2012}.

H. Liang, K. Wang and F. Zhu, "Mining social ties beyond homophily," 2016 IEEE 32nd International Conference on Data Engineering (ICDE), Helsinki, 2016, pp. 421-432.
\url{http://doi.org/10.1109/ICDE.2016.7498259} very datamining approach, they
look for link pattern with support and confidence above some threshold, that
also diverge from homophily by excluding attributes taking the same value (so
it's not exactly like us once again)

The kind of graphs with talked in this paper is a special case of what is
called \emph{heterogenous information network} in \autocite{HINSurvey17}.

learn a function that take as input the embedding of two nodes and output the predicted strength of
the relationship~\autocite{EdgeAwareEmbedding17}, could it be modified to output a cluster index?
\fi

>>>>

\Textcite{Taskar04} represent a node-attributed graphs with partially labelled edges as a
probabilistic graphical model (called Relational Markov Model) and learns its parameters from the
data using gradient descent. It is a flexible modeling approach that let the designer express
complex graph patterns. Using a slightly less principled approach, \textcite{conceptualLinks12}
first cluster nodes based solely on their attributes and then count the links between such groups to
keep only the maximally frequent ones, arguing this reveals the set of attributes that support the
connection between two nodes. In the same data mining vein, and related to our departure from global
homophily, \textcite{beyondHome16} look for link patterns whose support and confidence is not only
above some given thresholds, but that also diverge from homophily. In the author explicitly exclude
attributes taking the same value within such link patterns.
The objective of \textcite{ZhangModelFree16} is eventually to hard cluster the nodes into
communities but their work similarities with ours. Namely, they tweak the traditional modularity
objective (maximizing the density of intra community edges) by weighting the edges with the
similarity of their endpoints, and simultaneously learning a weight vector $w_\ell$ for a community
$C_\ell$. Formally they defined the $R$ criterion as :
\begin{equation*}
  R(\mathcal{C}, \beta;\alpha, \mu) = \sum_{\ell=1}^k \frac{1}{|\mathcal{C}_\ell|^\alpha}
  \sum_{u,v \in \mathcal{C}_\ell} A_{uv} \mu -e^{-g(s_{uv}, w_\ell)}
\end{equation*}
where $A$ is the adjacency matrix, $\alpha$ penalizes unbalanced communities, $\mu$ is a trade off
parameter between information from edges and node features, and $g$ is exactly our goodness
function.  They optimize it by alternately optimizing over the labels with fixed parameters and over
the parameters with fixed labels, using block coordinate descent.
Finally we mention a more quantitative work. \Textcite{Abraham2012a} assume that there are $K$
social categories modelled by $K$ Euclidean spaces $\mathcal{D}_i$. Nodes of the graph have an
associated point in each of these spaces. The key assumption is that space have small local
correlation: informally, the intersection of any two small balls from two distinct spaces is small.
These $K$ spaces give rise to small world networks $\mathcal{G}_i$, where the edge probability
$\propto \mathcal{D}_i(u, v)^{-d}$ and we observe the real network $\mathcal{G} =\bigcup_i
\mathcal{G}_i$. From $\mathcal{G}$, the proposed algorithm recovers in $n \mathrm{polylog} n$ time a
bounded approximation $\mathcal{D}'_i$ of all $\mathcal{D}_i$, that is
\begin{equation*}
  \sigma \mathcal{D}_i(u, v) \leq \mathcal{D}'_i(u, v) \leq \delta \mathcal{D}_i(u, v) + \Delta
\end{equation*}

@@
